---
title: "Synthetic Data Code Demo"
author: "Joseph Allen"
date: "8 June 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Import packages
```{r import packages, warning=FALSE}
library(readr)
library(knitr)
library(rmarkdown)
library(synthpop)
```


## Load dataset
```{r load temp.csv, warning=FALSE}
df <- read_csv("NATSAL/natsal_3_teaching_open_with_personal.csv")

# print table
# kable(head(df))
paged_table(df)
```


## Snynthpop
Synthpop has a codebook function which will assess your data, this runs before synthesis anyway but it can be useful to pre-process the data before synthesis.


``` {r synthpop-codebook, layout="l-body-outset",  warning=FALSE}
codebook.syn(df)
```


It's suggested that no column has more than 60 unique factors, for this reason we will drop our personal data, and reduce scope.
``` {r drop-cols, layout="l-body-outset",  warning=TRUE}
# Synthesis will fail due to more than 60 levels of data in our factors
# Drop columns with too many factors, and reduce
drops <- c("first_name","last_name","email", 'importance_religion', 'age_at_first_child')
df <- df[ , !(names(df) %in% drops)]

# print table
paged_table(df)
```

All strings need to be converted into factors
``` {r convert-strings, layout="l-body-outset",  warning=TRUE}
# convert strings to factors
df <- as.data.frame(unclass(df), stringsAsFactors = TRUE)

codebook.syn(df)
```

Finally we can synthesize a dataset
``` {r synthesis-time, layout="l-screen",  warning=FALSE}
# R Synthesis, synthesize 5 datasets
syn_df <- syn(df, m = 5)

# summarise
summary(syn_df)
```

We can compare our dataset with our new synthetic dataset, we can also compare with multiple synthetic datasets
``` {r compare, layout="l-body-outset",  warning=FALSE}
# We can then compare these, with multiple datasets
# compare(syn_df, df, msel = 1:5)
compare(syn_df, df, vars = "age_group", msel = 1:5)
```


Finally we can output our dataset to a CSV for our Jupyter analysis.
``` {r write-data, layout="l-body-outset",  warning=FALSE}
# synthpop has it's own write function that supports SPSS, CSV and much more

write.syn(syn_df,"syn_NATSAL", filetype = "csv", convert.factors = "numeric")
```

